Identifying Inter-Genomic Repeats Using Fast, Approximate Methods for Betweenness Centrality

### 2015-05-18 manuscript/manuscript.tex

Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path (SSSP) problem for each node. An SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.   The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation. Brandes algorithm calculates $\sigma_{st}$  simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$. In a subsequent aggregation phase, nodes are procesed by non-decreasing distance from $s$. So we have: 
$$\delta_{s}(v) = \sum_{w \in succ(v)} \frac{\sigma_{sv}}{\sigma_{sw}}(1+\delta_{s}(w))$$

where $\delta_{s}(v) = \frac{\sigma_{st}(v)}{\sigma_{st}}$ and $succ(v)$ is an immediate successor of $v$  in the shortest path DAG.   
This algorithm takes time $\Theta(mn)$ for unit edge weight graphs and $\Theta(mn + n^{2}log(n))$ for weighted graphs. This algorithm does not scale well with large networks with millions of nodes and edges. Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm. However this approach scales to about a few million nodes and may not be fast enough for larger assembly graphs.

>>>>

Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single-source shortest path (SSSP) problem for each node. A SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$. The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation. Brandes algorithm calculates $\sigma_{st}$  simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$. In a subsequent aggregation phase, nodes are processed by non-decreasing distance from $s$. So we have: 
$$\delta_{s}(v) = \sum_{w \in succ(v)} \frac{\sigma_{sv}}{\sigma_{sw}}(1+\delta_{s}(w))$$

where $\delta_{s}(v) = \frac{\sigma_{st}(v)}{\sigma_{st}}$ and $succ(v)$ is an immediate successor of $v$  in the shortest path DAG.   
This algorithm takes time $\Theta(mn)$ for unit edge weight graphs and $\Theta(mn + n^{2}log(n))$ for weighted graphs. This algorithm does not scale well with large networks with millions of nodes and edges. Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm, but scales in practice to only a few million nodes, limiting its application to larger metagenomic assembly graphs that exceed million of nodes.
